# Helpers functions

These are some helper functions that are needed for the MCMC.

These functions are not documented for now. (TODO: do this).

<!-- send_to to put all helper functions into one file.. -->

```{r, send_to = 'R/all-helpers.R'}
#' A helper function to print the progress of a simulation. Place directly at
#' the beginning of the loop, before any computation happens.
#' @param isim isim.
#' @param nsim number of sim.
#' @param type type of job you're running. Defaults to "simulation".
#' @param lapsetime lapsed time, in seconds (by default).
#' @param lapsetimeunit "second".
#' @param start.time Start time, usually obtained using \code{Sys.time()}
#' @param fill Whether or not to fill the line.
#' @param beep Whether to beep when done.
printprogress <- function(isim, nsim, type="simulation", lapsetime=NULL,
                          lapsetimeunit="seconds", start.time=NULL,
                          fill=FALSE, beep=FALSE){

    ## If lapse time is present, then use it
    if(fill) cat(fill=TRUE)
    if(is.null(lapsetime) & is.null(start.time)){
            cat("\r", type, " ", isim, "out of", nsim)
    } else {
        if(!is.null(start.time)){
          if(isim == 1){
            lapsetime = 0
            remainingtime = "unknown"
            endtime = "unknown time"
          } else {
            lapsetime = round(difftime(Sys.time(), start.time,
                                       units = "secs"), 0)
            remainingtime = round(lapsetime * (nsim - (isim - 1)) / (isim - 1), 0)
            endtime = strftime((Sys.time() + remainingtime))
          }
        }
        cat("\r", type, " ", isim, "out of", nsim, "with lapsed time",
            lapsetime, lapsetimeunit, "and remaining time", remainingtime,
            lapsetimeunit, "and will finish at", endtime, ".")
        if(beep & isim==nsim){beepr::beep()}
    }
    if(fill) cat(fill=TRUE)
}

#' Helper function to print progress.
#' 
#' @param iter Current iteration.
#' @param Nmc Total number of MCMC iterations.
#'
#' @export
progress <- function(iter, Nmc){
    cat("MCMC iteration: ", iter,
        ".", round(iter/Nmc,3)*100, "% is done", fill = TRUE)
}

wcrossprod.fast <- function(x, w.sqrt, weighting = TRUE){
    if(weighting){
        x <- x * w.sqrt
    }
  return(Rfast::Crossprod(x,x))
}

#' 
self.crossprod <- function(x){ ## this function can be deleted. 
    if(ncol(x) == 1){
        ret <- x %*% t(x)
    }else{
        ret <- crossprod(x)
    }
  return(ret)
}

#' 
logsumexp <- function (x) {
  y = max(x)
  y + log(sum(exp(x - y)))
}

#' 
softmax <- function (x) {
  exp(x - logsumexp(x))
}

#'
softmax2 <- function(x){
    return(exp(x)/sum(exp(x)))
}

#'
logMNdensity <- function(y,invSig,logdet){
    ## The term "-log(2*pi)*d/2" is omitted
    return(- Rfast::Crossprod(y, invSig) %*% y /2 - logdet/2)
}

#'
KL.MN <- function(p,q){
    sum(p*log(p/q))
}

#'
pgdraw.mod <- function(b,c){
    if(b>0){
        ret <- pgdraw::pgdraw(b,c)
    }else{
##        print("in pgdraw: b=0")
        ret <- 0
    }
    return(ret)
}

#' Helper to change stick-breaking vector |pi.sb| to regular probabilities.
#' @param pi.sb Stickbreaking probability vector.
#' 
#' @return same sized vector of size K.
#' 
#' @export
SB2MN <- function(pi.sb){
    pi.sb <- c(pi.sb,1) ## append 1
    K <- length(pi.sb)
    pi.mn <- rep(0,K)
    pi.mn[1] <- pi.sb[1]
    for(kk in 2:K){
        pi.mn[kk] <- pi.sb[kk]*(1-sum(pi.mn))
    }
    return(pi.mn)
}

#' Helper to change stick-breaking vector |pi.sb| to regular probabilities. The
#' opposite function of SB2MN.

#' @param pi.mn Probability vector.
#'
#' @return same sized vector of size K.
#'
#' @export
MN2SB <- function(pi.mn){
    K <- length(pi.mn)
    pi.sb <- rep(0,K-1)
    pi.cs <- cumsum(pi.mn[-K])
    pi.sb[1] <- pi.mn[1]
    pi.sb[2:(K-1)] <- pi.mn[2:(K-1)]/(1-pi.cs[-(K-1)])
    return(pi.sb)
}

#'
gamma2pi <- function(gamma,Xp){
    XpGamma <- Rfast::Crossprod(gamma, Xp) ## K-1 x T
    pi.sb <- 1/(1+exp(-XpGamma))
    return(pi.sb)
}

#'
dtaMatCensor <- function(x,Cbox){
    d <- dim(x)[2]
    for(ii in 1:d){
        x[,ii] <- pmin(pmax(x[,ii],Cbox[ii,1]),Cbox[ii,2])
    }
    return(x)
}

#'
censorIndicator <- function(x,Cbox){
    d <- dim(Cbox)[1]
    if(dim(x)[2]!=d & dim(x)[1]==d){
        x <- t(x)
    }
    for(ii in 1:d){
        x[,ii] <- (x[,ii]==Cbox[ii,2]) - (x[,ii]==Cbox[ii,1])
    }
    return(x)
}

#'
sample.region <- function(cc,Cbox){
    limits <- NULL
    d <- length(cc)
    for(i in 1:d){
        if(cc[i]==-1){
            limits <- rbind(limits, c(-Inf,Cbox[i,1]))
        }else if(cc[i]==0){
            limits <- rbind(limits, c(-Inf,Inf))
        }else{
            limits <- rbind(limits, c(Cbox[i,2], Inf))
        }
    }
    colnames(limits) <- c("lower","upper")
    return(limits)
}


#'

rmatnorm.fast <- function(M,U,V){ 
    ## M is d x p
    ## U is d x d
    ## V is p x p 
    M <- as.matrix(M)
    U <- as.matrix(U)
    V <- as.matrix(V)
    d <- nrow(M) ## dim of U 
    p <- ncol(M) ## dim of V 
    chol.U <- Rfast::cholesky(U) ## upper triangular matrix 
    chol.V <- Rfast::cholesky(V)
    x <- Rfast::matrnorm(d, p)
    return(Rfast::Crossprod(chol.U, x) %*% chol.V + M)
}
    



#'
samp.trunc.normal <- function(yy, zz, cc.info, bounds.info, mu.mat, Sigma.ell, dimdat){
    cc <- abs(cc.info)==1
    nc <- cc==0
    d <- length(cc.info)
    lower.info <- bounds.info[1:d]
    upper.info <- bounds.info[(d+1):(2*d)]
    mu <- mu.mat[,zz]

    ## If all dimensions are censored.
    if(all(cc)){
      sig = Sigma.ell[,,zz]
      if(dimdat==1) sig = as.matrix(sig)
        yy <- tmvnsim::tmvnsim(nsamp=1,d,lower=lower.info,upper=upper.info,
                      mean= mu, sigma = sig)$samp

    ## If /not/ all dimensions are censored.
    } else {
      ## Conditional mean and variance
        slope <- Sigma.ell[cc,nc,zz] %*% solve(Sigma.ell[nc,nc,zz])
        cond.sig <- Sigma.ell[cc,cc,zz] - slope %*% Sigma.ell[nc,cc,zz]
        cond.m <- mu[cc]+slope %*% (yy[nc]-mu[nc])

      ## Sample from this conditional truncated normal
      if(dimdat==1) cond.sig = as.matrix(cond.sig)
      stopifnot("matrix" %in% class(cond.sig))
      imputed_y = tmvnsim::tmvnsim(nsamp = 1, length(cond.m),
              lower=lower.info[cc], upper = upper.info[cc],
              mean= cond.m, sigma = cond.sig)$samp
        yy[cc] <- imputed_y
    }
    if(!is.numeric(yy)) browser()

    return(yy)
}

#' 
impute.censored <- function(ww, yy, zz, cc.info.mat, bounds.mat,
                            mu.mat, Sigma.ell, dimdat){
  ## Setup
  stopifnot("array" %in% class(Sigma.ell))
    if(length(zz) == 0){
        ret <- NULL
    }else if (length(zz)==1){
        ret <- samp.trunc.normal(yy, zz, cc.info.mat, bounds.mat,
                                 mu.mat, Sigma.ell, dimdat)
    } else {
        ret <- t(sapply(1:length(zz), function(ii)
            samp.trunc.normal(yy[ii,], zz[ii],
                              cc.info.mat[ii,], bounds.mat[,ii],
                              mu.mat, Sigma.ell, dimdat)))
    }
    return(ret)
}


#'
gen.syn.dta <- function(T, K, p, d=3, Cbox=NULL, Pi =NULL, avg.clust.size=100){
    ## this function can be deleted 
    ### by default, d = 3
    nt <- stats::rpois(n=T,lambda=avg.clust.size*K)
    X <- t(sapply(1:p, function(pp) stats::arima.sim(list(order=c(1,0,0), ar=pp/(pp+1)/2), n=T)))
    Beta <- array(stats::rnorm(K*p*d), c(d,p,K))
    beta0 <- matrix(stats::rnorm(K*d), nrow=d,ncol=K)

    if(is.null(Pi)){
        Gamma0 <- matrix(stats::rnorm(K*p), nrow= p, ncol = K)
        Gamma0 <- t(apply(Gamma0,1,function(x) x-x[K]))/10
        gamma0 <- (1:K)/1
        Pi <- t(apply(t(X) %*% Gamma0+gamma0, 1, function(x) exp(x)/sum(exp(x))))
    }

    Sigma <- stats::rWishart(K, 1, diag(d))

    Z.list <- lapply(1:T, function(tt) sample(1:K, nt[[tt]], prob = Pi[tt,],replace = TRUE))

    Y.list <- NULL
    for(t in 1:T){
        Z <- Z.list[[t]]
        Y.list[[t]] <- t(sapply(Z,function(z)
          MASS::mvrnorm(n=1, mu = beta0[,z]+Beta[,,z]%*%X[,t],Sigma = Sigma[,,z])))
    }
    Ytrue.list <- Y.list

    if(is.null(Cbox)==FALSE){
        Y.list <- lapply(Y.list, function(x) dtaMatCensor(x,Cbox))
    }
    return(list(Y.list=Y.list, X=X, Z.list=Z.list, nt=nt, Cbox=Cbox, K=K,
                Ytrue.list = Ytrue.list,
                Pi = Pi, beta0 = beta0, Beta = Beta, Sigma = Sigma))
}


#'
loglik_eval <- function(mu.list, chol.Sig.list,
                        W.list, X.list, Y.list, Z.list,
                        nt.list, simple = FALSE, n.cores = 1){

    ## logPiZ <- mcmapply(function(xx,yy,mm){sapply(1:K, function(kk)
    ##     mvnfast::dmvn(yy, mm[,kk], chol.Sig.list[[kk]],log=TRUE,isChol = TRUE)) },
    ##     xx=X.list, yy = Y.list, mm=mu.list, mc.cores = min(n.cores, T),
    ##     SIMPLIFY = FALSE)

    if(simple == TRUE){
        ll.vec <- sapply(1:nt.list[[1]], function(id)
            mvnfast::dmvn(Y.list[[1]][id,],
                          mu.list[[1]][,Z.list[[1]][id]],
                          chol.Sig.list[[Z.list[[1]][id]]],
                          log=TRUE, isChol = TRUE))
        out <- sum(ll.vec * W.list[[1]]) / nt.list[[1]]
    }else{
        ll.sums  <-  parallel::mcmapply(function(ww, xx,yy,mm,zz,nt){
            sum(ww * sapply(1:nt,function(id)
                mvnfast::dmvn(yy[id,], mm[,zz[id]], chol.Sig.list[[zz[id]]],
                              log=TRUE,isChol = TRUE)) )},
            ww = W.list, xx = X.list, yy = Y.list,
            mm=mu.list, zz = Z.list, nt = nt.list,
            mc.cores = min(n.cores, T), SIMPLIFY = TRUE)
        out <- Reduce("+",ll.sums)/sum(unlist(nt.list))
    }
    return(out)
}
```


Here's a function that makes a censoring box (dimdat by 2 matrix, where the
first column is the lower cutoff, and the second column is the upper cutoff).
 
```{r}
#' Helper to obtain the censoring limits |Cbox|, which is a 2-colum matrix with columns |bounds.lower| and |bounds.upper|.
#'
#' @param ylist Data (list of d-column matrices) including censored particles.
#'
#' @return |dimdat| by 2 matrix.
#' @export
get_Cbox <- function(ylist){

  ## Setup
  dimdat = ncol(ylist[[1]])

  ## Define censor limits
  bounds.lower <- Rfast::rowMins(matrix(unlist(
      lapply(ylist, function(xx) Rfast::colMins(xx,value = TRUE))), nrow = dimdat), value = TRUE)
  bounds.upper <- Rfast::rowMaxs(matrix(unlist(
      lapply(ylist, function(xx) Rfast::colMaxs(xx,value = TRUE))), nrow = dimdat), value = TRUE)

  ## bounds.lower <- lapply(ylist,function(xx)
  ##     Rfast::colMins(xx,value = TRUE)) %>%
  ##     do.call(rbind,.) %>% Rfast::colMins(.,value = TRUE)
  ## bounds.upper <- lapply(ylist,function(xx)
  ##     Rfast::colMaxs(xx,value = TRUE)) %>%
  ##     do.call(rbind,.) %>% Rfast::colMaxs(.,value = TRUE)

  ## Modify censorship slightly
  Cbox <- cbind(bounds.lower, bounds.upper)
}
```

## Ball constraint on $X^{(t)} \beta_k$

This function seems to have a bug. 

```{r}
#' Obtain the g parameter by simulation. 
#' 
#' @param X the design matrix, TT by p, (that doesn't include the intercept) 
#' @param dimdat the dimension of the cytogram space 
#' @param maxdev Maximum deviation of cluster means away from its grand mean. 
#' @param numclust the number of experts 
#' @param ggvec the vector of g parameter values to calculate the prior probability by Monte Carlo samples  
#' @param Nmc Monte Carlo simulation sample size, with default value being 1e4. 
#' @param n.cores the number of CPU cores to be used for parallelization 
#' @param viz show the plot of the fitted relationship between the g parameter and the prior probability.
#'
#' 
#' @return the g parameter with desired prior probability on maxdev 
#'
#' @export 
maxdev_to_gg <- function(X, dimdat, maxdev, numclust, ggvec,
                           Nmc = 1e4 , prior.prob = 0.99, 
                         viz = FALSE, n.cores = 1, verbose = FALSE){
  ## Basic setup
  p = ncol(X) ## Remember, X is T x p 
  d = dimdat
  
  ## Helper function
  ## @param tX is the transpose of X
  ball.deviance <- function(gg, rr, tX, 
                            Nmc=5000,
                            nu0=d,
                            nu1=p+1,
                            S0=diag(d),
                            S1=diag(p+1)){
      inv.XTX  <- solve(tcrossprod(tX))
      max.deviance <- rep(NA,Nmc) 
      
      Sig.ell <- matrixsampling::rinvwishart(Nmc,nu0+d,S0)
      beta.ell <- apply(Sig.ell,3,function(xx)
          as.matrix(matrixNormal::rmatnorm(M = matrix(0,d,p), 
                                           U = xx, 
                                           V = inv.XTX*gg, 
                                           tol = .Machine$double.eps^0.5)),
          simplify = FALSE)
      xb <- lapply(beta.ell, function(bb) bb%*% tX) 
                                      # length of list: Nmc 
                                      # each element: d x T 
      max.deviance <- lapply(xb, function(mat){
          apply(mat,2,function(cols) crossprod(cols)) %>% max()
      }) %>% unlist()
      prob <- mean(max.deviance <= rr^2)  
      return(list(gg = gg, rr = rr, prob = prob))
  }
  
  gglist <- as.list(ggvec)##as.list(1:40/100)
  n.cores <- min(n.cores, length(gglist))
  if(verbose) cat("Calculating gg value numerically", fill=TRUE)
    plist <- parallel::mclapply(1:length(gglist), function(igg){
      if(verbose) printprogress(igg, length(gglist), "Candidate values")
      gg = gglist[[igg]]
      ball.deviance(gg, maxdev, t(X), Nmc = Nmc)$prob
    },
  mc.cores = n.cores)
  if(verbose) cat(fill=TRUE)

  ## Make linear interpolation at a fine grid
  res = stats::approx(x = unlist(gglist), y = unlist(plist), method="linear", n = 100)
  newx =  res$x
  newy = res$y
  
  ## Get closest point
  imin =  which.min(abs(newy - prior.prob ^(1/numclust))) 
    
  ## Make some plots to confirm
  if(viz){
      plot(gglist,plist,type = "l",
           ylab = "Prob outside of \n radius r", xlab = "Value of |gg|")
      graphics::abline(h=0.05,lwd=2,col = "red")
      graphics::abline(h=0.01,lwd=2,col = "red")
      ## lines(y=newy, x=newx,lwd=2,col="blue")
      ## abline(v=newx[imin])
  }
  
  return(newx[imin])
}
```


```{r}
#' Closed form for obtaining the g parameter by simulation. Currently only works for dimdat=1
#' 
#' @param X the design matrix, TT by p, (that doesn't include the intercept) 
#' @param dimdat the dimension of the cytogram space 
#' @param maxdev Maximum deviation of cluster means away from its grand mean. 
#' @param numclust the number of experts 
#' @param prior_prob prior probability
#' @param viz show the plot of the fitted relationship between the g parameter and the prior probability.
#'
#' 
#' @return the g parameter with desired prior probability on maxdev 
#'
#' @export 
maxdev_to_gg_closed_form <- function(X, dimdat, maxdev, numclust, prior_prob = 0.99, viz = FALSE){

  stopifnot(dimdat == 1)
  obj = svd(X)
  pp = ncol(X)
  TT = nrow(X)
  ## maxdev = 0.115525

  ## z_cutoff = 2.58 ## for 99% for one cluster
  z_cutoff = qnorm(1-(1-prior_prob^(1/numclust))/2)
  
  ## Form the constant c and matrix P, multiplied to form the prior covariance cP.
  P = obj$u %*% t(obj$u)
  c = (maxdev^2 / z_cutoff^2 / diag(P)) %>% median
  A = c * obj$v %*% diag(1/(obj$d^2)) %*% t(obj$v) 
  cP = (X %*% A %*% t(X)) ##%>% diag() %>% plot()

  return(c)
}
``` 

Testing out this function.


```{r}
set.seed(100)
X = rnorm(1000,0,1) %>% matrix(ncol = 10, nrow = 100)
obj = svd(X)
pp = ncol(X)
TT = nrow(X)
maxdev = 0.5
numclust = 1
prior_probs = seq(from = 0.1, to = 1, by = 0.05)
sapply(prior_probs, function(prior_prob){
  z_cutoff = qnorm(1-(1-prior_prob^(1/numclust))/2)
  
  ## Form the constant c and matrix P, multiplied to form the prior covariance cP.
  P = obj$u %*% t(obj$u)
  c = (maxdev^2 / z_cutoff^2 / diag(P)) %>% median()
  ## c = (maxdev^2 / z_cutoff^2 / mean(diag(P)))
  A = c * obj$v %*% diag(1/(obj$d^2)) %*% t(obj$v) 
  cP = (X %*% A %*% t(X)) ##%>% diag() %>% plot()
  
  ## Generate beta and form w = X beta_k
  many_betas = MASS::mvrnorm(100000, mu = rep(0, pp), Sigma = A)
  w =  X %*% t(many_betas) 
  
  ## Plotting
  if(FALSE){
    ## Make a plot of all w = X beta_k
    w %>% .[,1:500] %>% matplot(pch = 16, col = rgb(0,0,0,0.1))
    
    ## Draw the lines where the 95'th quantiles are.
    qs = w %>% apply(1, function(one_w){
      quantile(one_w, probs=0.975)})
    lines(qs, col = 'red')
    
    qs = w %>% apply(1, function(one_w){
      quantile(one_w, probs=0.025)})
    lines(qs, col = 'red')
    
    abline(h = maxdev, lwd = 2, lty = 2, col = 'red')
    abline(h = -maxdev, lwd = 2, lty = 2, col = 'red')
    text(x=50, y = maxdev*1.1, labels = "Maximum deviation")
    legend("topleft", col = 'red', lwd=1, lty=1, legend = c("0.95 quantile"))
  }
  sum((abs(w)<maxdev))/length(w)
}) -> coverages
plot(x = prior_probs, y = coverages, type = 'o', ylim = c(0,1), xlim = c(0,1),
     xlab = "Target prior probability of deviation\n of |w| = |X beta| above maxdev",
     ylab = "Empirical probability of deviation")
abline(0,1, col = 'red')
```


Here is the 3d version (not properly tested yet, doesn't seem to work).

```{r, eval = FALSE}
maxdev_to_gg_closed_form_3d <- function(X, dimdat, maxdev, numclust,
                                        prior_prob = 0.99, viz = FALSE){

  stopifnot(dimdat == 3)
  obj = svd(X)
  pp = ncol(X)
  TT = nrow(X)
  ## maxdev = 0.115525

  ## z_cutoff = 2.58 ## for 99% for one cluster
  ## z_cutoff = qnorm(1-(1-prior_prob^(1/numclust))/2)
  chisq_cutoff = qchisq(p=1-(1-prior_prob^(1/numclust)), df = dimdat)
  
  ## Form the constant c and matrix P, multiplied to form the prior covariance cP.
  P = obj$u %*% t(obj$u)
  c = (maxdev^2 / chisq_cutoff^2 /diag(P)) %>% median()
  A = c * obj$v %*% diag(1/(obj$d^2)) %*% t(obj$v) 
  cP = (X %*% A %*% t(X)) ##%>% diag() %>% plot()

  if(viz){
  
    ## Generate beta and form w = X beta_k
    ## many_betas = MASS::mvrnorm(100000, mu = rep(0, pp), Sigma = A)
    many_betas = matrixsampling::rmatrixnormal(n=100000,
                                               M=matrix(0, pp, dimdat),
                                               U=A,
                                               V = diag(rep(1,dimdat)))
    wlist = lapply(1:100000, function(ii)X %*% many_betas[,,ii])
    wlist_rownorms = wlist %>% lapply(function(w) apply(w, 1,function(a)sum(a*a)))
    
    ## Make a plot of all |w|_2, w = X beta_k
    ## wlist_rownorms %>% lapply(function(a)a[1]) %>%  .[1:500] %>% matplot(pch = 16, col = rgb(0,0,0,0.1))
    wlist_rownorms %>% .[1:500] %>% do.call(rbind, .) %>%  matplot(pch = 16, col = rgb(0,0,0,0.1))

    ## Draw the lines where the 95'th quantiles are.
    qs =     wlist_rownorms %>% .[1:500] %>% do.call(rbind, .) %>% apply(1, function(one_w){
      quantile(one_w, probs=0.95)})
    lines(qs, col = 'red')
    abline(h = maxdev, lwd = 2, lty = 2, col = 'red')
    text(x=50, y = maxdev*1.1, labels = "Maximum deviation")
    legend("topleft", col = 'red', lwd=1, lty=1, legend = c("0.95 quantile"))
  }

  return(c)
}
```



(This should reduce to the dimdat=1 case. Work in progress!)

```{r, eval = FALSE} 
set.seed(100)
X = rnorm(1000,0,1) %>% matrix(ncol = 10, nrow = 100)
obj = svd(X)
pp = ncol(X)
TT = nrow(X)
maxdev = 0.5
dimdat = 1
numclust = 1
prior_probs = seq(from = 0.1, to = 1, by = 0.1)
sapply(prior_probs, function(prior_prob){
  print(prior_prob)
  chisq_cutoff = qchisq(p = (1-(1-prior_prob^(1/numclust))), df = dimdat)
  ## print(1-(1-prior_prob^(1/numclust)))
  ## print(chisq_cutoff)

  ## Form the constant c and matrix P, multiplied to form the prior covariance cP.
  P = obj$u %*% t(obj$u)
  c = (maxdev^2 / chisq_cutoff^2 / diag(P)) %>% median()
  A = c * obj$v %*% diag(1/(obj$d^2)) %*% t(obj$v) 
  cP = (X %*% A %*% t(X)) ##%>% diag() %>% plot()
  
  ## Generate beta and form w = X beta_k
  ## many_betas = MASS::mvrnorm(100000, mu = rep(0, pp), Sigma = A)
  nsim = 10000
  many_betas = matrixsampling::rmatrixnormal(n = nsim,
                                             M = matrix(0, pp, dimdat),
                                             U = A,
                                             V = diag(rep(1,dimdat)))
  wlist = lapply(1:nsim, function(ii)X %*% many_betas[,,ii])
  wlist_rownorms = wlist %>% lapply(function(w) apply(w, 1,function(a)sum(a*a)))
  
  
  ## Make a plot of all |w|_2, w = X beta_k
  if(FALSE){
    ## wlist_rownorms %>% lapply(function(a)a[1]) %>%  .[1:500] %>% matplot(pch = 16, col = rgb(0,0,0,0.1))
    wlist_rownorms %>% .[1:500] %>% do.call(rbind, .) %>%  matplot(pch = 16, col = rgb(0,0,0,0.1))

    ## Draw the lines where the 95'th quantiles are.
    qs = wlist_rownorms %>% .[1:500] %>% do.call(rbind, .) %>% apply(1, function(one_w){
      quantile(one_w, probs=0.95)})
    lines(qs, col = 'red')
    abline(h = maxdev, lwd = 2, lty = 2, col = 'red')
    text(x=50, y = maxdev*1.1, labels = "Maximum deviation")
    legend("topleft", col = 'red', lwd=1, lty=1, legend = c("0.95 quantile"))
  }

  normmat = do.call(rbind, wlist_rownorms)
  hist(normmat[,1])
  abline(v = maxdev, col = 'red', lwd=2)

  sum((abs(normmat)< maxdev))/(length(normmat))
}) -> coverages
plot(x = prior_probs, y = coverages, type = 'o', ylim = c(0,1), xlim = c(0,1),
     xlab = "Target prior probability of deviation\n of |w| = |X beta| above maxdev",
     ylab = "Empirical probability of deviation")
abline(0,1, col = 'red')
```
```



Here's a function to take the stick-breaking $\gamma$ parameters (with $X$) and
transform them to a probability vector.

```{r}
#' Description goes here
#' 
#' @param ga document this carefully
#' @param Xp 1 vector appended to left of covariate matrix X.
#' 
#' @export
#' @return (OO x OO) matrix with each row..
post_process_pie <- function(Xp,ga){
  Xga <- Xp %*% ga 
  pie.SB <- 1/(1+exp(-Xga))
  return(t(apply(pie.SB,1, SB2MN)))
}
```



Here's a helper to change the results from an MCMC to a flowmix- (or flowtrend-)
like object. This way, we can use the plotting helpers from those packages.


```{r}
#' Reformatting the results from MCMC to create a flowmix-like object that
#' contains of the posterior means of the model parametr. This allows you to use
#' plotting code from the flowmix and flowtrend package.
#' 
#' @param res The result of doing MCMC.
#' @param last_draws_inds The last few MCMC draws to take. Defaults to NULL.
#'
#' @return posterior mean estimates of the cluster coefficients, mean, variance
#'   and probability.
#' 
#' @export
#' 
mcmc_res_to_flowmix <- function(res, last_draws_inds=NULL){

  ## Setup
  X = t(res$dat.info$X)
  numclust = res$pos.Sigma %>% .[3]
  TT = nrow(X)
  p = ncol(X)
  numclust = res$dat.info$numclust
  dimdat = res$dat.info$ylist[[1]] %>% ncol()
  obj = list()
  Nmc = res$pos.beta %>% dim() %>% tail(1)
  if(is.null(last_draws_inds)) last_draws_inds = 1:Nmc
  if(!is.null(last_draws_inds))stopifnot(all(last_draws_inds<=Nmc))
  ## last_draw_inds = 1:dim(res$pos.beta)[4]

  ## Means
  pos.mn <- list()
  for(kk in 1:numclust){
    last_draws_inds
    pos.mn[[kk]] <- mclapply(last_draws_inds, function(mm){
      res$pos.beta[,,kk,mm,drop=TRUE]%*% rbind(1,t(X))
    },
    mc.cores = detectCores())%>%
      abind::abind(.,along=3)
  }
  post.mn.mean <- lapply(pos.mn, function(aa){
    apply(aa,c(1,2),mean)})
  obj$mn = array(NA, dim = c(TT, dimdat,numclust))
  for(iclust in 1:numclust){
    obj$mn[,,iclust] = t(post.mn.mean[[iclust]]) 
  }

  ## probabilities
  ## pos.gamma.mean = res$pos.gamma[,,last_draws_inds] %>% apply(c(1:2), mean)
  pos.SB <- apply(res$pos.gamma[,,last_draws_inds], c(2,3), function(ga)
    1/(1+exp(-t(ga) %*% rbind(1,t(X)))))
  pos.MN <- apply(pos.SB, c(1,3), flowcut:::SB2MN)  
  post.pi.mean <- apply(pos.MN,c(1,2), mean)
  obj$prob = t(post.pi.mean)

  ## cluster covariances
  obj$sigma = array(NA, dim = c(numclust, dimdat, dimdat))
  obj$numclust = numclust
  res$pos.Sigma %>% dim()
  ## post.Sigma.mean <- apply(res$pos.Sigma,c(1,2,3), mean) %>% aperm(c(3,1,2))
  pos.Sigma.mean = res$pos.Sigma[,,,last_draws_inds] %>% apply(c(1,2,3), mean) %>% aperm(c(3,1,2)) 
  obj$sigma = pos.Sigma.mean

  ## mean regression coefficients
  pos.beta = res$pos.beta[,,,last_draws_inds]
  pos.beta.mean = pos.beta %>% apply(c(1,2,3), mean) 
  obj$beta = pos.beta.mean
  
  ## prob regression coefficients
  pos.gamma.mean = res$pos.gamma[,,last_draws_inds] %>% apply(c(1:2), mean)
  obj$alpha = pos.gamma.mean

  ## Bundle and return
  obj$numclust = numclust
  obj$TT = TT
  class(obj) = "flowmix"
  return(obj)
}
```

Here's a function to impute the y's based on a model defined by the posterior
mean of the last 500 draws.


```{r}
#' Imputing particles to impute the y's based on a model defined by the posterior
#' mean of the last 500 draws. This is separated out from the Gibbs since the result takes up a lot of memory.
#' (This is all code in the Gibbs.fast() function.)
#'
#' 
#' @param ylist Original particles
#' @param countslist Biomass list
#' @param Cbox Bounding box.
#' @param mcres The converted MCMC result, using e.g., |mcmc_res_to_flowmix(res,
#'   last_draws_inds = 1001:1500)|
#' @param n.cores number of CPU cores to use
#' 
#' @export
impute <- function(ylist, countslist, Cbox, mcres, n.cores=1){

  ## if(FALSE){
  ##   ## Temporary
  ##   ylist = datobj$ylist
  ##   countslist = datobj$countslist

  ##   ## TEMPORARY: Add a bounding box |Cbox|
  ##   bounds.lower <- lapply(datobj$ylist,function(xx)
  ##     Rfast::colMins(xx,value = TRUE)) %>%
  ##     do.call(rbind,.) %>% Rfast::colMins(.,value = TRUE)
  ##   bounds.upper <- lapply(datobj$ylist,function(xx)
  ##     Rfast::colMaxs(xx,value = TRUE)) %>%
  ##     do.call(rbind,.) %>% Rfast::colMaxs(.,value = TRUE)
  ##   Cbox <- cbind(bounds.lower,bounds.upper)
  ##   Cbox <- res$dat.info$Cbox
  ##   datobj$Cbox = Cbox

  ##   ## Temporary
  ##   mcres = mcmc_res_to_flowmix(res, last_draws_inds = 1001:1500)
  ## }
  
  ## Get everything
  ntlist = sapply(ylist, nrow)
  TT = length(ylist)
  NN <- sum(ntlist)
  dimdat <- ncol(ylist[[1]])
  countsTotal <- sapply(countslist,sum) %>% sum() 
  alpha.factor <- NN/countsTotal  
  W.list <- lapply(countslist, function(xx) xx*alpha.factor) 

  ## More processing
  Censor.list <- lapply(ylist, function(x) censorIndicator(x,Cbox))  
  censor.01.list <- lapply(Censor.list, function(x) apply(x,1, function(xx){
    sum(abs(xx)) > 0})) ## expensive
  censor.which.list <- lapply(censor.01.list, function(cc) which(cc==TRUE))
  censored.ylist <- mapply(function(yy,c01) {yy[c01==TRUE,,drop=FALSE]},
                                       yy = ylist, c01 = censor.01.list)
  censored.C.list <- mapply(function(c01,cc){cc[c01==TRUE,,drop=FALSE]},
                                        c01 = censor.01.list, cc=Censor.list)
  censored.W.list <- mapply(function(c01,ww){ww[c01==TRUE]},
                                        c01 = censor.01.list, w=W.list)
  ntlist.censor <- sapply(censor.01.list, sum) 

  Z.list = flowmix::gate(mcres, ylist, NULL, 1111, eps_estep = 1E-20)

  censored.Z.list <- mapply(function(c01,zz){zz[c01 == TRUE]},
                            c01 = censor.01.list, zz = Z.list)

  ## Sort of expensive
  samp.region.list <- lapply(1:TT, function(tt){
      c01 <- censor.01.list[[tt]]
      cc <- Censor.list[[tt]]
      if(sum(c01==TRUE)>1){
          apply(cc[c01==TRUE,,drop=FALSE],1,function(xx)
              ## flowcut:::sample.region(xx,Cbox))
              sample.region(xx,Cbox))
      }else if(sum(c01==TRUE)==1){
          matrix(sample.region(cc[c01==TRUE,,drop=FALSE],Cbox),ncol=1)
      }else{
          NULL
      }
  })

  ## Get the censored particles
  Sig.ell = mcres$sigma %>% aperm(c(2,3,1))
  imputed.ylist <- parallel::mclapply(1:TT, function(tt) {
    mu.mat = mcres$mn[tt,,]
    yy = impute.censored(ww = censored.W.list[[tt]], 
                         yy = censored.ylist[[tt]],
                         zz = censored.Z.list[[tt]],
                         cc.info.mat =  censored.C.list[[tt]],
                         bounds.mat = samp.region.list[[tt]],
                         ##mu.list[[tt]]
                         mu.mat = mu.mat, Sigma.ell = Sig.ell,
                         dimdat = dimdat)
    if(dimdat==1 & !is.null(yy)) yy = t(yy)
    return(yy)
  }, mc.cores = n.cores, mc.preschedule = FALSE)             


  ## Finally, ca
  imputed_ylist <-
    parallel::mcmapply(function(yy, c01, imp){replace(yy, c01, imp)}, 
                       yy = ylist, c01 = censor.01.list, imp = imputed.ylist,
                       mc.cores = n.cores, SIMPLIFY = FALSE)            
  return(list(imputed_ylist = imputed_ylist, censor.01.list = censor.01.list,
              memlist = Z.list))
}
```

