# Simulations

Refer back to the data generation in \@ref{syntheticdata}. We will be
top-censoring the data at 0.5. Before censoring, this is what the data looks
like.

```{r, fig.width = 6, fig.height = 4}
## Load the "original" model
orig_model = readRDS(file=file.path("~/repos/flowcut/inst/output", "orig_model.RDS"))

## Generate data
set.seed(100)
isignal = 8
new_model = make_model(orig_model, isignal)
ylist = gen_1d(new_model, nt = 100)
flowtrend::plot_1d(ylist, obj = new_model)
```

After censoring, this is it.


```{r, fig.width = 6, fig.height = 4}
## Censor it
ylist = lapply(ylist, function(y){
  y = pmin(y, 0.5)
})
flowtrend::plot_1d(ylist, obj = new_model)
```

We need to specify a few things (1) like the censoring limits `Cbox` and (2)
`countslist` before running the MCMC.

```{r}
## Form the censoring "box"
Cbox = rbind(c(-Inf, 0.5)) 

## Counts are all equal for now
countslist = lapply(ylist, function(y){ rep(1, nrow(y)) })

## Save the metadata
datobj = list(ylist=ylist, countslist=countslist, Cbox=Cbox, X=orig_model$X)
saveRDS(datobj,
        file.path("~/repos/flowcut/inst/output",
                  paste0("isignal-", isignal, "-datobj.RDS")))
```

We also need some *prior elicitation* to prevent the cluster means from changing
too much across time.

(code copy-pasted as-is for now, not meant to be run)

```{r, eval = FALSE}
##############################
###### prior elicitation
##############################

d <- 3
p <- 39 
X <- t(readRDS("pre-censor-datobj.RDS")$X)
    
nu0=d
nu1=p+1 
S0=diag(d)
S1=diag(p+1)
inv.XTX  <- solve(tcrossprod(X))
dim(inv.XTX)

## Sig.ell <- rinvwishart(1,nu0+d,S0)[,,1]
## gg <- 1 
## beta.ell <- rmatnorm(M = matrix(0,d,p), 
##                      U = Sig.ell, 
##                      V = inv.XTX*gg, 
##                      tol = .Machine$double.eps^0.5)
## mean.vec <- apply(beta.ell%*% X,1,mean)
## l2.norm <- mean(apply(beta.ell%*% X,2,function(xx)
##     crossprod(xx-mean.vec)))
## hist(apply(beta.ell%*% X,2,function(xx)
##     crossprod(xx-mean.vec)),breaks = "FD")

ball.deviance <- function(gg,rr,X, Nmc=5000,
                          nu0=d, nu1=p+1,
                          S0=diag(d), S1=diag(p+1),
                          simple= TRUE){
    inv.XTX  <- solve(tcrossprod(X))
    ms.deviance <- rep(NA,Nmc) 

    Sig.ell <- rinvwishart(Nmc,nu0+d,S0)
    beta.ell <- apply(Sig.ell,3,function(xx)
        as.matrix(rmatnorm(M = matrix(0,d,p), 
                 U = xx, 
                 V = inv.XTX*gg, 
                 tol = .Machine$double.eps^0.5)),
        simplify = FALSE)
    xb <- lapply(beta.ell, function(bb) bb%*% X)  
    mean.vec <- lapply(xb, function(bb)
        apply(bb,1,mean))
    ms.deviance <- mapply(function(xx,mm){
        mean(apply(xx,2,function(cols) crossprod(cols-mm)))
    }, xx=xb,mm=mean.vec)
    prob <- mean(ms.deviance > rr^2)
    if(simple){
        return(list(gg=gg,rr=rr,prob=prob))
    }else{
        return(list(gg = gg, rr = rr,
                    prob=prob,
                    msd = ms.deviance)) 
    }
}

gglist <- as.list(1:100/100)
plist <- mclapply(gglist,function(gg)
    ball.deviance(gg,0.5,X,Nmc=1e4)$prob,
    mc.cores = min(n.cores, length(gglist)))
plot(gglist,plist,type = "l") 
abline(h=0.05,lwd=2,col = "red")
abline(h=0.01,lwd=2,col = "red")
lines(gglist,lowess(plist~gglist,f=0.3)$y,lwd=2,col="blue")




hist(ball.deviance(0.01,0.5,X)$msd, breaks = "FD")
hist(ball.deviance(0.1,0.5,X)$msd, breaks = "FD")
hist(ball.deviance(0.2,0.5,X)$msd, breaks = "FD")
hist(ball.deviance(0.5,0.5,X)$msd, breaks = "FD")
hist(ball.deviance(1,0.5,X)$msd, breaks = "FD")
hist(ball.deviance(2,0.5,X)$msd, breaks = "FD")
```


Next, we run the MCMC.

```{r, eval = FALSE}
## Run the MCMC
Nmc <- 1e3 * 5
Nburn <- 500
Gibbs.res <- run.Gibbs.fast(ylist = datobj$ylist,
                            countslist = datobj$countslist,
                            numclust = 2,
                            Nmc = Nmc, Nburn = Nburn,
                            gg = 0.1, 
                            X = t(datobj$X),
                            Cbox = datobj$Cbox, verbose = TRUE)

## Save the results
saveRDS(Gibbs.res,
        file.path("~/repos/flowcut/inst/output",
                  paste0("isignal-", isignal, "-gibbs.RDS")))
```
