# Simulations

Refer back to the data generation in \@ref{syntheticdata}. We will be
top-censoring the data at 0.5. Before censoring, this is what the data looks
like.

```{r load-data2, fig.width = 6, fig.height = 4, eval = FALSE}
## Load the "original" model
orig_model = readRDS(file=file.path("~/repos/flowcut/inst/output", "orig_model.RDS"))

## Generate data
set.seed(100)
isignal = 8
new_model = flowcut::make_model(orig_model, isignal)
ylist = flowcut::gen_1d(new_model, nt = 100)
flowtrend::plot_1d(ylist, obj = new_model)
```

After censoring, this is it.


```{r, fig.width = 6, fig.height = 4, eval = FALSE}
## Censor it
ylist = lapply(ylist, function(y){
  y = pmin(y, 0.5)
})
flowtrend::plot_1d(ylist, obj = new_model)
```

We need to specify a few things (1) like the censoring limits `Cbox` and (2)
`countslist` before running the MCMC.

```{r, eval = FALSE}
## Form the censoring "box"
Cbox = rbind(c(-Inf, 0.5)) 

## Counts are all equal for now
countslist = lapply(ylist, function(y){ rep(1, nrow(y)) })

## Save the metadata
datobj = list(ylist=ylist, countslist=countslist, Cbox=Cbox, X=orig_model$X)
saveRDS(datobj,
        file.path("~/repos/flowcut/inst/output",
                  paste0("isignal-", isignal, "-datobj.RDS")))
```

We also need some *prior elicitation* to prevent the cluster means from changing
too much across time.

(code copy-pasted as-is for now, not meant to be run)

```{r, eval = FALSE}
gg <- maxdev_to_gg(datobj$X, 
                   dimdat = 3,
                   maxdev = 0.5,
                   numclust = 10,
                   ggvec = (1:40)/200, 
                   n.cores = detectCores(),
                   Nmc = 1e4, viz=FALSE)
gg <- 0.01287879
## hist(ball.deviance(0.01,0.5,t(X))$msd, breaks = "FD")
## hist(ball.deviance(0.1,0.5,X)$msd, breaks = "FD")
## hist(ball.deviance(0.2,0.5,X)$msd, breaks = "FD")
## hist(ball.deviance(0.5,0.5,X)$msd, breaks = "FD")
## hist(ball.deviance(1,0.5,X)$msd, breaks = "FD")
## hist(ball.deviance(2,0.5,X)$msd, breaks = "FD")
```


Next, we run the MCMC.

```{r, eval = FALSE}
## Run the MCMC
Nmc <- 1e3 * 5
Nburn <- 500

Nmc <- 20
Nburn <- 10

set.seed(123)
Gibbs.res <- run.Gibbs.fast(ylist = datobj$ylist,
                            countslist = datobj$countslist,
                            numclust = 2,
                            Nmc = Nmc, Nburn = Nburn,
                            gg = 0.1, 
                            X = t(datobj$X),
                            Cbox = datobj$Cbox, verbose = TRUE)

## Save the results
saveRDS(Gibbs.res,
        file.path("~/repos/flowcut/inst/output",
                  paste0("isignal-", isignal, "-gibbs.RDS")))
```
