# Real data example

We run the Gibbs sampler on real data.

We'll load the data, process it, then feed it into the MCMC.

TODO: bundle the Cbox into a `make_Cbox()` function. Why are we doing
`more.censor()` manually? Simplify/streamline this.

```{r, eval = FALSE}
## Load the code for running MCMC
## codedir =  "~/Dropbox/Apps/Overleaf/censor-flowmix/BMEcensor/code/"
codedir =  "~/Dropbox/Apps/Overleaf/censor-flowmix/BMEcensor/code/"
source(file.path(codedir, 'BME-functions.r'))

## Load dataset
datadir = "~/Dropbox/research-new/censored-flowmix/code/data"
dta <- readRDS(file.path(datadir, "pre-censor-datobj.RDS"))

## Assign data to objects
ylist <- dta$ylist
TT <- length(ylist)
d <- 3
countslist = dta$countslist

## Define censor limits
bounds.lower <- rowMins(matrix(unlist(
    lapply(dta$ylist, function(xx) colMins(xx,value = TRUE))),
    nrow = d), value = TRUE) 
bounds.upper <- rowMaxs(matrix(unlist(
    lapply(dta$ylist, function(xx) colMaxs(xx,value = TRUE))),
    nrow = d), value = TRUE)

## Censor the first dimension a bit more
Cbox <- cbind(signif(log(bounds.lower),5), log(bounds.upper))
more.censor <- function(ymat,bd){
    cbind(sapply(ymat[,1], function(xx) ifelse(xx<bd, bd, xx)),
          ymat[,c(2,3)]) }
ylist <- mclapply(dta$ylist,
                  function(yy){ more.censor(log(yy),Cbox[1,1]), },
                  mc.cores = min(n.cores, TT))


## Checking that censoring is done properly
assertthat::assert_that(all(rowMins(matrix(unlist(
                lapply(ylist, function(xx) colMins(xx,value = TRUE))),
                nrow = d), value = TRUE) == Cbox[,1]))
assertthat::assert_that(all(rowMaxs(matrix(unlist(
                lapply(ylist, function(xx) colMaxs(xx,value = TRUE))),
                nrow = d), value = TRUE)  == Cbox[,2]))


## Clean data
datobj <- list(ylist = ylist,
               countslist = dta$countslist,
               X = dta$X,
               numclust = 10,
               Cbox=Cbox) 


## Run the MCMC
Nmc <- 1e3*5
Nburn <- 500
K  <-  10
Gibbs.res0 <- run.Gibbs.fast(ylist = datobj$ylist,
                             countslist = datobj$countslist,
                             X = datobj$X,
                             numclust = datobj$numclust,
                             gg = 0.1,
                             Nmc = Nmc, Nburn = Nburn,
                             Cbox = datobj$Cbox, verbose = TRUE,
                             warm.start = NULL)
```





# Documenting the package and building

We finish by running commands that will document, build, and install the
package.  It may also be a good idea to check the package from within this file.

```{r, results='hide'}
litr::document() # <-- use instead of devtools::document() 
```
